Model to study co-evolution of trade an culture

### 2015-03-16 article.tex
start to modify everything for state of the art trade

\section{PREVIOUS MODELISATIONS OF TRADE AND CULTURE}

@@

\xrc{I would suggest to use modelling instead of modelisation}
\subsection{Modelisation of Trade}

Multiple approaches have been proposed to model the economical interactions between humans. Among them, this section will review three which constitute major trends and are the most relevant to our study. The approaches are presented in chronological order.

\xrc{I'm not sure if it would be wiser to limit this section to models of ancient trade...otherwise a proper lit. review becomes huge.}

The first approach to be reviewed is the Expected Utility (EU) which became a main trend after the second world war~\cite{schoemaker_expected_1982}. The goal of this model was to propose a first explanation to the economical actions of humans along the following axes: descriptive, predictive, postdictive and prescriptive. More precisely this approach model individuals as agents facing a choice among risky options. Each agent is aware of the probability that an event will happen and the utility (economical reward) associated to that event. Each agent can then maximise its expected utility by choosing the proper actions. By applying this methodology to various economical problems it became clear that humans do not see problems as probabilities and do not directly reason on them. From the perspective of this theory, humans show a collection of irrational behaviours, notably a strong aversion to risk.

In an attempt to overcome the shortcomings of the EU models a new theory was introduced, the prospective theory. This theory conserve an approach close to the EU: individuals are modelled as agents facing risky actions each with a certain probability and utility. However, the agents do not face these problems by maximising the expected utility. They are rather endowed with behaviours composed of three key elements~\cite{camerer_prospect_2004}. First, the ``loss aversion'' element models the fact that individuals are much more sensitive to loss than to gains of the same magnitude. Second, the ``reference dependence'' models the fact that individuals measure their losses and gains with regards to their current possessions. Finally, individuals do not weight linearly the probabilities e.g. an individual will perceive a higher gain between 100 and 200 than between 1000 and 1100. Multiple difficulties have been encountered in applying this approach to practical problems of the real life. These are attributed to the lack of definition of certain elements of the theory and therefore the near impossibility to measure them~\cite{barberis_thirty_2012}.

The two previous approaches address an economical problem as a block which all agents face in a similar fashion. As a consequence, the emerging properties due to the interaction of numerous agents can not be studied. Moreover, these approaches are not suitable to model systems composed of individuals with various characteristics and interests. The Agent-based Computational Economics (ACE) fields propose specifically to model the complex outcomes raising from a large number of interactions in a decentralized market~\cite{tesfatsion_agent-based_2003}. Naturally this approach is interesting for our project based on multi-agent methodology and it constitutes a source of hypothesis to test to characterise the economical network of the Roman empire. More specifically, multiple hypothesis are proposed on the formation of different types of economical networks depending on the economical rules between the agents~\cite{kirman_evolving_2001,tesfatsion_structure_2001}. Another interesting track is the study of the evolution of behavioural norms in an economical network~\cite{epstein_learning_2001,axelrod_complexity_1997}. Finally~\cite{gintis_emergence_2006} has proposed a specific analysis on the emergence of price structure when only pair exchanges are considered. 
\todo{put the section}
This latest work constitutes naturally a strong reference and will be explained in further details in section XXX.

>>>>

\section{PREVIOUS MODELLINGS OF TRADE AND CULTURE}

@@

\subsection{Modelling of Trade}

Multiple approaches have been proposed to model the economical interactions in the ancient mediteranean culture. Among these we will focus on the formal models. These are close to our works, can be tested and compared.

%Hopkins 1980: structure of tax system. two hypothesis. 1)taxation increased volume of trade in the empire? 2) tax-exporting province (recently integrated province) had to export to pay tax. roman supported by many small taxes. Not sure to cite it, goes to much more
%Termin 2001: based on multiple consideration classify roman economy as market based with loose tie from things far appart. composed of local markets connected together
%Terpstra 2011: micro economic model to explain possibility of trade in roman empire under uncertainty. based on reputation in the network and removal of bad from the network (personalized network based on geographical info). focus on a specific case (one city and its relations with others). threat of expulsion or loss of reputation as reinforcement <- not only economic.

Finally we would like to add this additional work even if it has not been specifically designed for historical studies~\cite{gintis_emergence_2006}. This model proposes a specific analysis on the emergence of price structure when only pair exchanges are considered, thus observing the emergence of a market economy. This constitutes a relatively realistic model with simple assumptions and will be explained in further details in section XXX.
